
##### Basis functional model


## Use 3 different basis 
# W1 : truncated power basis
# W2 : polynomial radial basis
# W3 : B-Spline basis
```{r}
n = 500
i = 1:n
xi = (2*i - 1)/1000
```

## Approximate the target function and compare the column space of design matrix
```{r}
k = 3
L = 11
H = L + k + 1
xh = seq(min(xi), max(xi), length.out = L + 2)[2:(L+1)]  # knots
```

### Generate design matrix 
```{r}
## Design matrix W1 and W2 of using truncated power basis and using polynomial radial basis, respectively.
W1 = matrix(rep(0, n*H), ncol=H)  # Design matrix of using truncated power basis
W2 = matrix(rep(NA, n*H), ncol=H) # Design matrix of using polynomial radial basis
for(j in 1:L){
  W1[, j] = (xi - xh[j])^3
  W2[, j] = (abs(xi - xh[j]))^3
}
W1[W1<0] = 0

for(j in (L+1):H){
  W1[, j] = W2[, j]= xi^(j-L-1)
}

W3 = bs(xi, knots=xh, degree=3, intercept=T)  # Design matrix of using B-Spline basis
```

### Check whether the three design matrices span the same column space.
```{r}
all.equal(W1%*%solve(t(W1)%*%W1)%*%t(W1), W2%*%solve(t(W2)%*%W2)%*%t(W2))
all.equal(W1%*%solve(t(W1)%*%W1)%*%t(W1), W3%*%solve(t(W3)%*%W3)%*%t(W3))
all.equal(W2%*%solve(t(W2)%*%W2)%*%t(W2), W3%*%solve(t(W3)%*%W3)%*%t(W3))

(W1%*%solve(t(W1)%*%W1)%*%t(W1))[1:5, 1:5]
(W2%*%solve(t(W2)%*%W2)%*%t(W2))[1:5, 1:5]
(W3%*%solve(t(W3)%*%W3)%*%t(W3))[1:5, 1:5]
```

### Generate $y_i$ 
```{r}
set.seed(512)
y = sin(2*pi*xi^3)^3 + rnorm(n, 0, 0.1)
plot(xi, y, cex=0.5)
```

### Approximate the target function with three basis
```{r}
fit1 = lm(y ~ W1)
fit2 = lm(y ~ W2)
fit3 = lm(y ~ W3)
pred1 = predict(fit1)
pred2 = predict(fit2)
pred3 = predict(fit3)

mfrow=c(1,3)
plot(xi, y, cex=0.5)
lines(xi, pred1, col="red", lwd=1.2)
plot(xi, y, cex=0.5)
lines(xi, pred2, col="red", lwd=1.2)
plot(xi, y, cex=0.5)
lines(xi, pred3, col="red", lwd=1.2)
```




## Use g-prior and L~Poison(1)

### Assign a prior for parameters.
g-prior on the $\beta_H$ with $g=n$ i.e.$\beta_H | \sigma^2, H, y \sim \text{N}_H \left({g \over 1+g}(W_H^TW_H)^{-1}W_H^Ty,\, {g\sigma^2 \over 1+g}(W_H^TW_H)^{-1}\right)$
$p(\sigma^2) \propto \sigma^{-2}$
$L\sim \text{Pois}(1)$


### Obtain the model averaged pointwise posteior for $\mu(x_0)$
```{r}
n = 500
j = 1:999
x0 = j/1000
k = 3
g = n

L_temp = seq(1,40)       # Set the range of L

# Compute the posterior probability of H
prob.L=c()
for(h in L_temp){
    xh = seq(min(xi), max(xi), length.out = h+2)[2:(h+1)]  # knots
    WH = bs(xi, knots=xh, degree=k, intercept=T)
    prob.L[h] = dpois(h, 1, log=T) - n/2*log(0.5 * t(y) %*% (diag(rep(1, n)) - g/(1+g)*WH%*%solve(t(WH)%*%WH)%*%t(WH)) %*%y)
}

# Normalize the posterior probability using log-sum-exp trick
log_prob_sum = max(prob.L) + log(sum(exp(prob.L - max(prob.L))))
prob.L = exp(prob.L - log_prob_sum)
```

```{r}
niter = 10000

betaH_samp = matrix(rep(NA, niter*(max(L_temp)+k+1)), nrow=niter)
mu_x = matrix(rep(NA, niter*length(x0)), nrow=niter)
L_samp = c()
news2 = c()
for(t in 1:niter){
  # Sample H and construct B-spline basis
  newL = sample(L_temp, 1, prob=prob.L)
  newH = newL + k + 1
  xh = seq(min(xi), max(xi), length.out = newL+2)[2:(newL+1)]  # knots
  newWH = bs(xi, knots=xh, degree=k, intercept=T)
  
  inv_WH = solve(t(newWH)%*%newWH)
  
  # Sample sigma2 from the posterior distribution given H
  news2[t] = rinvgamma(1, n/2, 0.5 * t(y) %*% (diag(rep(1, n)) - (g/(1+g))*newWH%*%inv_WH%*%t(newWH)) %*%y)
  
  # Sample beta_H from the posterior distribution given H and sigma2
  betaH = mvrnorm(1, g/(1+g)*inv_WH%*%t(newWH)%*%y, g/(1+g)*news2[t]*inv_WH)
  
  # mu
  xh = seq(min(x0), max(x0), length.out = newL+2)[2:(newL+1)]
  WH = bs(x0, knots=xh, degree=k, intercept=T)
  for(i in 1:length(x0)){
    mu_x[t, i] = sum(betaH*WH[i, ])
  }
  
  # Store the parameters
  for(h in 1:H){
    betaH_samp[t, h] =  betaH[h]
  }
  L_samp[t] = newL
  
}
```

### Trace plot of L and sigma2
```{r}
plot(news2, type="l")
plot(betaH_samp[,2], type="l")
```

```{r}
par(mfrow=c(1,2))
hist(L_samp, main="Histogram of Sample of L")
hist(news2, main="Histogram of Sample of sigma2")

table(L_samp)
```


### Posterior mean and credible interval
```{r}
post_mean_mux = apply(mu_x, 2, mean)
post_LB_mux = apply(mu_x, 2, quantile, 0.025)
post_UB_mux = apply(mu_x, 2, quantile, 0.975)
```

### Draw the posterior mean and the 95% credible interval for every $x_0$
```{r}
ggplot(mapping = aes(x=x0, y=post_mean_mux)) +
  geom_line(mapping = aes(x=xi, y=sin(2*pi*xi^3)^3), color="red", size=1.0, alpha=0.8) +
  geom_line(color="blue", size=1.0, linetype="dashed") +
  geom_ribbon(aes(ymin=post_LB_mux, ymax=post_UB_mux), alpha=0.4) +
  geom_point(mapping=aes(x=xi, y=y), alpha=0.3, size=0.2)+
  ggtitle("Posterior mean and the 95% credible interval") 

```



